Scaling up a GF grammar and prepare it for cool stuff for Swedish

### 2011-11-17 doc/report/report.tex
passive and FocAdv working
reflexive, updates to bara, formal object

\begin{document}
% ska gå att förstå och locka till läsning
%ska tydligt beskriva vad rapporten handlar om
%ska innehålla vettiga sökord
\title{A large scale grammar and parser for Swedish}
\author{Malin Ahlberg}
\maketitle
\newpage

@@

\section{Background}  
Too long way to build a parser/grammar from scratch. Idea to combine existing parts. 
The project is heavily dependening on three resources, will describe them here.
\subsection{GF}
Most important part. Been developed for ? years, inspiration from Haskell, ML,
and also Prolog, Agda. Can be used with other programming langauges. 
Main ideas, abstract and concrete, translation, parse- and abstract trees.
What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
ideas of anaphores, montague? dependent types, ontologies.
For this project, only (mostly) important to model Swedish without abandoning the
resources too much. Use of the big grammar for translation, why it might be hard with
semantics and lexicon.
Research about how to parse with gf by Krasimir, experiments with English.
Example of a piece with Swedish, the reader gets familiar with how it looks.

@@

\subsection{The grammar}
Although we can't expect to get a full coverage, we would like a tillfredsställande
coverage of standard Swedish. May also be interesting for other purposes.
Explain what's in the resources, constructions present in most languages. 
Swedish shares with Nor and Dan. Extra module, covering things like sw is a
v2-lang, focusing parts of sentences. Sw also makes use of passive s-form a lot.
kommer att - for scand, extra tempus, example
passiv s: but now need bigger change in the grammar. 
embedded - relative
pronouns/Quant
Genitive -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form
           fixing this include restructuring all
Det-utr

@@

The VP require many changes, to add passive, PPart, bara.
About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
for the last one, we need VPSlash to allow adverb.

varandra and sin/sitt/sig. Not subject, but NP.

@@

\subsubsection{Testing}
Hard do test, can't verify. Regression testing. Talbanken is still too hard.
Test how many parse trees that are generated.
Some results.

@@

\section{Future Work}
What is still left to do? Description and direction of future work:
how to make the parser robust, elipses and long sentences, names and numbers.
What is still needed in the grammar, how important that is. 
how to get valency information, description of extracting thise.
make the parser faster by using smaller lexicon.
Lexicon tool on the fly.
Getting and using probabilities, to disambiguate.
process list from saldo.
Use multiligual treebank?

>>>>

\begin{document}
% ska gå att förstå och locka till läsning
%ska tydligt beskriva vad rapporten handlar om
%ska innehålla vettiga sökord
\title{Scaling up a GF grammar and prepare it for cool stuff for Swedish}
\author{Malin Ahlberg}
\maketitle
\newpage

@@

\section{Background}  
Too long way to build a parser/grammar from scratch. Idea to combine existing parts. 
The project is heavily dependening on three resources, will describe them here.
\subsection{GF}
Most important part. Been developed for ? years, inspiration from Haskell, ML,
and also Prolog, Agda. Can be used with other programming langauges. 
Main ideas, abstract and concrete, translation, parse- and abstract trees.
What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
ideas of anaphores, montague? dependent types, ontologies.
For this project, only (mostly) important to model Swedish without abandoning the
resources too much. Use of the big grammar for translation, why it might be hard with
semantics and lexicon.
Research about how to parse with gf by Krasimir, experiments with English.
Example of a piece with Swedish, the reader gets familiar with how it looks.
Also show how information 'disappears', become strings.

@@

\subsection{The grammar}
Although we can't expect to get a full coverage, we would like a tillfredsställande
coverage of standard Swedish. Been working on a grammar fragment based on Talbanken,
like cool people like Montague. May also be interesting for other purposes.
Explain what's in the resources, constructions present in most languages. 
Swedish shares with Nor and Dan. Extra module, covering things like sw is a
v2-lang, focusing parts of sentences. Sw also makes use of passive s-form a lot.
kommer att - for scand, extra tempus, example
passiv s: but now need bigger change in the grammar. 
embedded - relative
pronouns/Quant
Genitive -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form
           fixing this include restructuring all
Det-utr

@@

The VP require many changes, to add passive (no), PPart, bara.
About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
for the last one, we need VPSlash to allow adverb.

varandra and sin/sitt/sig. Not subject, but NP.
work pretty well by letting all subject depend on the subject in the phrase.
Extra info, but true 'han såg sin mammas bil', they same could be applied to
APs for 'han är längre än sin kompis' or Adv for 'han leker oftare än sin kompis'. (ComparAdvAdj)
The current version is nice and works for recursive genitivs, but atm you can't focus
this kind of object, then a ClSlash need info about the subject. FunRP also tricky, in this
case RP needs to give more information.

@@

\subsubsection{Testing}
Hard do test, can't verify. Regression testing. Talbanken is still too hard.
While developing: regression, 150 sentences to start with. Made longer, both good and bad,
to see how many ways things can be parsed.
Test how many parse trees that are generated.
Some results.

@@

\section{Future Work}
What is still left to do? Description and direction of future work:
how to make the parser robust, elipses and long sentences, names and numbers.
What is still needed in the grammar, how important that is. 
how to get valency information, description of extracting thise.
make the parser faster by using smaller lexicon.
Lexicon tool on the fly.
Getting and using probabilities, to disambiguate.
process list from saldo.
Use multiligual treebank?
sTop ambiguities by fixing so that Gen of a quant isn't a quant.
